TBD

### 2020-12-07 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Edits to section 2 on defining classical system and section 6.3 on Curiel's counterarguments

@@

Provide any relevant background on classical mechanics, specifically to clarify/define terminology. Might be able to just put this in the introduction proper, or the start of the next section. Make explicit how our terminology relates to other standard terminology in the philosophical literature.

A classical Hamiltonian system is any system whose phase space is a symplectic manifold and that evolves according to Hamiltonian evolution (i.e. according to a symplectomorphism).

Throughout, we will focus only on non-dissipative classical systems, i.e. conservative systems. For in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $. This requires that the system is conservative, which is equivalent to the system being deterministic and reversible. Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
% Dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential

@@

@@

>>>>

Like North \parencites*[85]{North}, our argument involves investigating Hamiltonian systems outside the hyper-regular domain.

@@

Goal of section: Provide any relevant background on classical mechanics, specifically to clarify/define terminology. Might be able to just put this in the introduction proper, or the start of the next section. Make explicit how our terminology relates to other standard terminology in the philosophical literature.

% [To do: figure out if North and/or Curiel restrict their discussion/argument to the hyper-regular domain]. It seems as though Jill North is explicitly not assuming that her argument applies only to the hyper-regular domain [page 85, especially footnote 50]. Part of her argument involves motivating the idea that we will not in general be able to construct an equivalent Lagrangian formulation, and this involves appealing to systems outside the hyper-regular domain.
% Figure out explicitly how North defines classical systems. It might be in terms of the equations of motion that hold for classical particles [87]. North is arguing that ``symplectic structure is a simpler, more natural, more general underlying structure" that ``suffices for" this theory of classical mechanics [87]. Seems similar to GC position.

Various classes of classical systems are commonly delimited by specifying a set of equations that they satisfy, such as Hamilton's equations or the Euler--Lagrange equations. We will instead delimit classical systems based on the physical conditions that they satisfy, such as reducibility and deterministic evolution. We will then derive the relevant equations as consequences of these physical assumptions. For instance, we will see that a classical Hamiltonian system is any system whose phase space is a symplectic manifold and that evolves according to Hamiltonian evolution (i.e. according to a symplectomorphism).

The assumption of reducibility plays the key role of distinguishing classical systems from quantum-mechanical ones. Unlike classical systems, quantum systems are \textit{irreducible}, in the sense that specifying the state of the whole system is not equivalent to specifying the states of its parts (and vice versa). In contrast, a reducible system comprises subsystems whose states jointly determine the state of the whole system. In the widest sense, a classical system is a reducible one.

Newtonian systems do not necessarily have deterministic and reversible evolution, for instance in the case of a damped harmonic oscillator. Yet, they do satisfy the motion equivalence assumption because knowledge of the position and velocity allows for the determination of the acceleration [why does this entail that motion equivalence is satisfied?]. In contrast, Hamiltonian systems are necessarily deterministic and reversible, although they need not satisfy motion equivalence. We will understand Lagrangian mechanics as the intersection of Newtonian and Hamiltonian systems: they have deterministic and reversible evolution, while also satisfying motion equivalence.

By focusing on systems with deterministic and reversible evolution, we necessarily exclude dissipative classical systems (i.e. ones that do not conserve energy). For determinism and reversibility are equivalent to a system being conservative. This restriction arises partly from the following reason: in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we will assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $. This requires that the system is conservative. Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations. We, of course, do not deny the sense in which the Lagrangian formulation applies to dissipative systems, but this is not what we will mean by ``Lagrangian mechanics" \parencites[]{Smith}{Curiel}.\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems (by moving to the extended phase space), although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime. We will view dissipative systems as Newtonian, but neither Lagrangian nor Hamiltonian systems [appeal to Norton's dome for motivation?]

Further motivation for focusing on non-dissipative systems: dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential. Or, as Curiel puts it, the systems involve a ``force $1 $-form that is not exact and that is not Lie-derived along the Lagrangian vector field solving the inhomogeneous system" [311].

% Question of how Newtonian mechanics fits and: we can understand Newtonian mechanics as a hybrid between Hamiltonian and Lagrangian mechanics. On the one hand, Newtonian systems need not be deterministic and reversible. On the other hand, they satisfy the motion equivalence assumption, because knowledge of the position and velocity allows for the determination of the acceleration.

@@

\url{https://physics.stackexchange.com/questions/398159/formulation-of-an-action-principle-in-a-general-phase-space-no-cotangent-bundle?rq=1}

seems to have very similar kind of argument

@@

Like our argument, part of Curiel's argument involves appealing to systems outside of the hyper-regular domain. Curiel acknowledges both that only hyper-regular Lagrangians have unique solutions and that Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}. Nevertheless, he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage. On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}. In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy a particular set of physically significant kinematical constraints (his equations 4.3). Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will still satisfy the kinematical constraints of Lagrangian mechanics \parencites*[311]{Curiel}.

Summary: Hamiltonian mechanics provide solutions outside hyper-regular domain (but not these kinematical constraints) whereas Lagrangian mechanics preserves these kinematical constraints outside hyper-regular domain. Worry: perhaps GC and Curiel are in a verbal dispute about the proper/privilege definition of a classical system (with Curiel privileging these kinematical constraints, and GC privileging deterministic/reversible evolution).

On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems). In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers Hamiltonian $H=p$, which corresponds to system moving with constant velocity (e.g. velocity of 1). So this could very well be a model of a photon, moving at the speed of light. Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating Euler--Lagrange equation. Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310]. But then does Curiel consider this system to not be a classical system?

Question to ask GC: does he agree with the following claims made by Curiel regarding the differences between the symplectic structure in Hamiltonian mechanics (which is fixed) vs. the symplectic structure that arises from the Euler--Lagrange equations (which varies from Lagrangian to Lagrangian): \\ 
Hamiltonian mechanics: ``in Hamiltonian mechanics, the fact that the symplectic structure is fixed once and for all enforces the satisfaction of the kinematical constraints (5.3), and it encodes the existence and uniqueness of solutions to the equations of motion and the conservation of energy, among many other things. All of these features are independent of the individual Hamiltonian used in the equation" [310]. \\
Lagrangian mechanics: ``In Lagrangian mechanics, in contrast, it is the Lagrangian itself that encodes existence and uniqueness of solutions (which will obtain when and only when the Lagrangian is regular), and there is not even a guarantee that energy is conserved, as one can, for example, represent dissipative systems in the Lagrangian framework" [310].

% Either way, presumably we can agree with Curiel's explanation/understanding for why the Lagrangian must be regular: Curiel arguing that he can give an account of ``why it makes physical sense to demand that the Lagrangian be regular", i.e. satisfy equation 4.6 for the Hessian determinant. He notes that by considering the variational problem, we understand ``why the Lagrangian must be quadratic (at least) in the velocities for there to be a unique solution to the Euler--Lagrange equation" [308]. Argument: for the variational problem to be well-posed, you need to be able to ``take non-trivial derivatives of the Lagrangian up to second-order in the velocity, since the crucial integration by parts that yields the Euler--Lagrange equation demands this, on pain of giving, for example, the tautology $0=0$" in some cases, or a contradiction in others.
